---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
unzip("./activity.zip")
activity <- read.csv("./activity.csv")
```


## What is mean total number of steps taken per day?

```{r}
actbydate <- tapply(activity$steps, activity$date, sum)
hist(actbydate, main="Total Steps Taken each Day", xlab="Number of Daily Steps")
meansteps <- mean(actbydate, na.rm = TRUE)
mediansteps <- median(actbydate,na.rm = TRUE)
```

The mean total steps taken each day is `r meansteps` and the median total steps taken each day is `r mediansteps`.

## What is the average daily activity pattern?
```{r}
library('dplyr')
actbyint<-activity %>% 
    group_by(interval)%>%
    summarise(steps=mean(steps,na.rm = TRUE))

plot(x=actbyint$interval, y=actbyint$steps, type = "l", main="Time Series of Steps",
     xlab='Time (in minutes)',ylab='Steps')
max <- max(actbyint)
maxint <- which(actbyint==max(actbyint))
```
The interval `r maxint` contains the maxium average number of steps at `r max` steps. 

## Imputing missing values

```{r}
amtNA <- sum(is.na(activity))
```

In the original dataset there are `r amtNA` missing data points.To remove these missing data points, I will replace the unknown values with the mean of the corresponding interval. 

```{r}

newactivity <- activity

for(i in 1:nrow(newactivity)){
    if(is.na(newactivity[i,1])){
        
        #Look up mean for interval from actbyint table from previous problem         and replace it.
        newactivity[i,1] = actbyint[which(actbyint$interval == newactivity[i,3]), 2]
    }
} 
newactbydate <- tapply(newactivity$steps, newactivity$date, sum)
hist(newactbydate, main="Total Steps Taken each Day", xlab="Updated Number of Daily Steps")
newmeansteps <- mean(newactbydate)
newmediansteps <- median(newactbydate)
```


After updating the data by replacing the missing data with the mean for the interval for the given missing data, the new mean is `r newmeansteps` steps and the new median is `r newmediansteps` steps.The mean for the original dataset and the updated data set is the same. The median is slightly lower for the updated data. Since I had used the average number of steps for a given interval for the missing data points, it did not change the average.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
newactivity$PartOfWeek <- NA
newactivity$date<-as.Date(newactivity$date)
for(i in 1:nrow(newactivity)){
    if(weekdays(newactivity[i,2])=='Sunday' || weekdays(newactivity[i,2])=='Saturday'){
        newactivity[i,4] = 'Weekend'
    }
    else{
        newactivity[i,4]='Weekday'
    }
} 


library('ggplot2')
ggplot(newactivity, aes(x=interval, y=steps))+
    stat_summary(fun.y=mean, geom='line')+facet_grid(~PartOfWeek)+
    ggtitle('Average Steps Taken')
```
